The Spatial and Temporal Domains of Modern Ecology

### 2017-12-20 paper/SI.tex
Finished editing the SI

The following text provides additional information on this study's Methods and Results.  The full manuscript, data, and code for all calculations are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}. \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript. In the interim, the entire project folder can be downloaded from the following \href{https://www.dropbox.com/sh/1l9em6p0wyabwr3/AADhCpFvqjv6y7VbL5tgB3Iqa?dl=0}{link}.}

\subsection*{Methods}% Make the title.
\vspace{-4pt}
%\maketitle 
\subsubsection*{Paper selection}
\vspace{-10pt}
We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.  To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Table S1) between 2004 and 2014. Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.  A separate set of 20 papers was also randomly selected, and this set was given to all observers to review independently, in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.

\centerline{
%\hspace{1cm} 
\parbox{11cm}{\noindent Table 1. The selected journals and their 2012 impact factors.}
}
\vspace{-12pt}
%\begin{longtable}[h]
\begin{longtable}{lr}
%\centering
%\caption{The selected journals and their 2012 impact factors.}\\
%\centering
%\begin{tabular}{lr}
  \hline
\textbf{Journal} & \textbf{Impact Factor} \\
\hline
\endhead 
  \hline
Ecology Letters & 17.95 \\ 
  Ecological Monographs & 8.09 \\ 
  Frontiers In Ecology And The Environment & 7.62 \\ 
  Global Ecology And Biogeography & 7.22 \\ 
  Global Change Biology & 6.91 \\ 
  Diversity And Distributions & 6.12 \\ 
  Methods In Ecology And Evolution & 5.92 \\ 
  Proceedings Of The Royal Society B-biological Sciences & 5.68 \\ 
  Journal Of Ecology & 5.43 \\ 
  Ecology & 5.17 \\ 
  Ecography & 5.12 \\ 
  Journal Of Biogeography & 4.86 \\ 
  Functional Ecology & 4.86 \\ 
  Journal Of Animal Ecology & 4.84 \\ 
  Journal Of Applied Ecology & 4.74 \\ 
  American Naturalist & 4.55 \\ 
  Conservation Biology & 4.36 \\ 
  Ecological Applications & 3.81 \\ 
  Biological Conservation & 3.79 \\ 
  Biogeosciences & 3.75 \\ 
  Bulletin Of The American Museum Of Natural History & 3.48 \\ 
  Biology Letters & 3.35 \\ 
  Oikos & 3.32 \\ 
  Behavioral Ecology & 3.22 \\ 
  Ecosystems & 3.17 \\ 
  Advances In Ecological Research & 3.08 \\ 
  Oecologia & 3.01 \\ 
  Landscape Ecology & 2.90 \\ 
  Agriculture Ecosystems \& Environment & 2.86 \\ 
  Ecological Economics & 2.85 \\ 
   \hline
%\end{tabular}
\end{longtable}
% Place your abstract within the special {sciabstract} environment.

@@

We recorded six primary dimensions of ecological observations, three related to space and three related to time.  The space-related dimensions were resolution, extent, and actual extent. Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was defined as the summed area of all sample plots (i.e. N * resolution, where N is the number of spatial replicates, which we also recorded), or the area that ecologists observe in practice. In assessing spatial scales, our analysis only considered the Cartesian plane. We did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g. depth profiles in marine ecology). In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.

@@

A full definition of all dimensions and how they were recorded is contained within the answers to the list of questions below. This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, in order to ensure methodological consistency (see next section).

@@

\noindent\underline{\textbf{General:}}
\def\Hitem{\item [Q\stepcounter{enumi}\arabic{enumi}.]} 
\def\Hsubitem{\item [Q\arabic{enumi}.\arabic{enumi}.]} 
\vspace{-0.5cm}
\begin{enumerate}
  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) book reviews; 3) model-only studies, particularly theoretical models, which are not developed or tested against observed data; 4) if they are experimental manipulations (but if there is a study that has a mix of observational and experimental, record the observational treatments and exclude the manipulated treatments).

  \Hitem \emph{\textbf{What are the standard categories to be used for defining “Study type”?}} Define study type according to the following categories: Remote sensing, passive/automated data collection, other geographic data (e.g. non-remotely sensed GIS data), field/direct observation, or paleo-reconstruction (tree rings, charcoal cores, etc).

@@

  \Hitem \emph{\textbf{What is sampling duration, and how do we record it?}} How long an individual observation of an individual point in space took to make. Sampling duration multiplied by the number of repeats observations is used to calculate study duration. Often this value will not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the sampling duration. For example, for a field based method with intensive plot methods, if you can't estimate a plausible sampling duration, assign a token 1 day. For remote sensing observations you can assume one second (although the observations are effectively instantaneous).  
   \Hitem \emph{\textbf{What is duration, and how do we estimate it?}} The duration is the total period of time over which the phenomenon of interest was observed. More specifically, in the case of repeated observations, this is the total time that elapsed between the first and last observations at a given point in space (or of the same individual organism or community). For once-off (unreplicated) observations, this time is equivalent to the sample duration. However, there may be cases where once-off observations may have a longer duration than the sample duration.  For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013}. The observations are therefore not strictly temporally replicated, but the authors control for year of collection in their subsequent analysis to avoid confounding effects.  In this case, we can consider the effective duration to be three years, as the temporal information is encoded in the analysis. 
  \Hitem \emph{\textbf{What is \emph{actual duration}, and how do we calculate it?}} The actual duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question. To clarify, actual duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (duration), nor the integral of time spent in observing all spatial replicates. 
  \Hitem \emph{\textbf{Should actual duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g. for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated above, actual duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
  \Hitem \emph{\textbf{How do we record duration and actual duration when there are no repeat observations?}} In these cases, study duration is equal to sampling duration. 
   \Hitem \emph{\textbf{How do I record interval in cases where the interval is inconsistent?  For example, in a study were observations were observations were repeated in 1979, 1980, 1981, 1984, 2007, 2009?}} Find the time between each successive period, and then take the average of that (remember to convert to days!). If there are two or more sets of unevenly spaced days for each site/plot/measurement being taken in the study, then find the average interval for each, and average the averages.  
    \Hitem \emph{\textbf{How do you determine the interval for paleo-reconstructions?}} Use the minimum estimate for dating precision as the estimate of time between samples (e.g. 50 years in the study of European charcoal deposits DOI:10.1111/geb.12090). 
    \Hitem \emph{\textbf{How do you determine the sample duration (our third time category) for paleo-reconstructions?}} Similar to the previous question, the sampling duration is also the same as the minimum estimate of dating precision. The logic behind this is that in such cases, where a sediment or tree core or similar measurement is being collected, this effectively represents a continuous ``observation'', and the value associated with the minimum (or other reported) interval is typically an average (or another summary statistic like the maximum) of the amount accumulated. 
    \Hitem \emph{\textbf{What about intervals and durations for instrument-collected, or automatically-collected, observations?}} These are often similar in essence to the paleo-reconstruction case. Take the minimum temperature or daily rainfall recorded at a weather station, which require constant observation over 24 hours to report. In such cases, the interval and sampling duration are both 24 hours. On the other hand, automated logging systems will provide a series of high frequency observations that are collected instantaneously. In these cases the sampling duration should be a token one second (to keep consistent with remote sensing [Q7]), and the interval should be the period between successive instantaneous measurements. 
    \Hitem \emph{\textbf{How do you treat interval for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably some season-specific phenomenon (e.g. breeding behavior), the reported values should be pegged to the season, not averaged across the duration (the start and end dates of the study). So in this example, it would be 10 days.

@@

%But I'm noticing now that this isn't updated to include how to handle when the study unit is animals. I guess we could just add a sentence saying that the effective extent is the surface area of an average-sized individual of the animal species, or something along those lines. Let me know what you think.

%  Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares. For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots. For spatially contiguous studies (e.g. those based on remote sensing imagery), it should be the total area covered by the imagery, i.e. pixel resolution multiplied by the number of pixels. However, only record the area analyzed by the authors. If they used a sub-sample of pixels, extent is the number of those pixels a multiplied by pixel resolution.

@@

To ensure consistency within the main analysis, the lead author also reviewed each observer's results from their individual draw of papers and flagged values that appeared to deviate from the protocol for re-review by the observer. Re-reviewed values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.

@@

We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 371 observations over 1000 iterations. For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting (from a uniform distribution) a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion. In certain cases, the perturbation resulted in values for some certain dimensions that were not physically possible (e.g. interval longer or actual duration longer than duration; actual extent larger than extent), in which case we adjusted the perturbed values for the dimension that should be the smallest to equal that of the one that should be largest (e.g. we set interval to equal the perturbed duration value). This perturbed set of observations provided a basis for estimating uncertainty within our extracted set of observational scales.

\subsubsection*{Scale metrics}
\vspace{-10pt}
In presenting results (Fig. 1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1000 perturbed ensemble members in order to account for the large range in scale values. We calculated histograms for the four primary observational dimensions from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig. 1 main text).

To reveal the densities of observations within two dimensions (Fig. 2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated. Bandwidths of varying resolutions were tested on kernel density estimates were tested on kernel density estimates of sampling interval versus plot resolution (see Supplementary Results section).

@@

\noindent Where x is either extent or duration and y is its actual counterpart. We then evaluated how the magnitudes of difference varied with increasing values of actual extent/duration, using box plots to summarize decades within the same bins used for extent and duration (Fig. 1B and D main text). Decades were calculated for each pair for all bootstrap replicates. The whiskers in the summary box plots (Fig. 3 main text) fall slightly below 0 in some bins, even though extent and duration can never be smaller than their actual value. This is an artifact of the perturbation applied to estimated values, which in some instances caused extent/duration values to fall below their actual value. We retained these discrepancies rather than forcing such cases to zero to avoid biasing the summary statistics.

\subsubsection*{Trends in methods and scale}
To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period. To undertake the former assessment, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.  For the second analysis, we applied the same regression analysis to the four primary dimensions (resolution, extent, interval, and duration) in order to assess whether there were any trends in observational scales.

@@

We assessed variability and consistency between observers along multiple dimensions. First, we assessed how reliably observers agreed with respect to selecting or rejecting papers for review, using the R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.

@@

Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table S2). CV values were estimated from all six observers for resolution, actual extent, and interval, from five observers (all but Choi) for duration, and from three observers (Elsen, Estes, and Treuer) for extent. The smaller numbers for duration and extent reflect that initial efforts were focused on estimating resolution, interval, and actual extent and actual duration, thus fewer observers were available to assess extent and extent.

\subsubsection*{Between-observer coefficient of variation}
\vspace{-10pt}
We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.

\begin{table}[ht]
\captionsetup{width=1\linewidth}
\caption{Estimates of variability calculated from each observer's estimates of the spatial and temporal scales of ecological observations reported within a common set of the 20 papers used for calibration. Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records. }
\vspace{-4 pt}
\begin{center}
\begin{tabular}{rR{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}}
\hline
& \multicolumn{3}{c}{\textbf{Spatial}} & \multicolumn{3}{c}{\textbf{Temporal}}\\\cline{2-7}
%Value & Resolution & Extent & \multirow{2}{*}{Actual} & Interval & Duration & Actual Duration \\
Value & Resolution & Extent & Actual Extent & Interval & Duration & Actual Duration \\
\hline
CV of overall mean & 50 & 50 & 78 & 49 & 36 & 109 \\
Mean of record-wise CV & 58 & 41 & 72 & 105 & 64 & 126 \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

%\begin{figure}[!ht]
%%\begin{wrapfigure}{c}{1\textwidth}
%\includegraphics[width=1\textwidth]{figures/res_v_extent_pbstrap2.pdf}
%\vspace{-0.15 cm}
%\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left) and temporal duration and spatial extent (right). Density estimates were applied to the log-transformed values of each observational dimension after being randomly perturbed within uncertainty bounds drawn from an assessment of between-observer variability. Density estimates are rescaled to represent percentages. }
%\label{afoto1}
%\end{figure}
%

@@

Fig. \ref{obshists} details the distributions of observations by method along the four primary dimensions, while Fig. \ref{obskde} reveals the densities of observations by method within the dimensional juxtapositions shown in Fig. 2 in the main text. Fig. \ref{obsdiff} shows the how the differences between actual extent and extent  and actual duration and duration vary by observational type.

\subsubsection*{Trends in methods and scales}
\vspace{-10pt}
The results of weighted linear regression applied to observational types by year (Fig. \ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (remote observation regression, p$<$0.12; field observation regression, p$<$0.18).  Automated sensing methods showed no visible trend.

If the trend towards increasing use of remote sensing between 2004-2014 was real and not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average.

Of course, we cannot attach confidence to a trend in observational extent extrapolated from a non-significant trend in observational type. However, some additional evidence in support of this possible trend lies within the extent values themselves. The regression analyses applied to dimensions shows an apparent increase in extent of 0.25 orders of magnitude per year (Fig. \ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This weak trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).

\subsubsection*{Choice of bandwidth in kernel density estimation}
\vspace{-10pt}
Figure \ref{ksens} indicates the effect that varying bandwidth has on the appearance of the kernel density estimates. The smallest bandwidth (0.4) tested (Fig. \ref{ksens} top) shows that the primary observational concentrations described in the main text are evident, but tend to be devolved into separate concentrations. An example is the oblong concentration of observations that in Fig. 1A (and in the lower left panel in Fig. \ref{ksens}) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions. With the smaller bandwidth this concentration appears as two separate patches (Fig. \ref{ksens} top left), but with 0.7 bandwidth applied becomes coherent (Fig. \ref{ksens} middle left).  This increasing separation of points into separate clusters is expected as bandwidth is reduced.  However, this assessment indicates that the choice of bandwidth does not alter the interpretation of the results, as the domains of highest and lowest concentrations remain the same.

>>>>

The following text provides additional information on this study's Methods and Results.  The full manuscript, data, and code for all calculations are available as part of an R package that can be installed from an accompanying \href{https://github.com/agroimpacts/ecoscales}{GitHub repository}. \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript. In the interim, the entire project folder can be downloaded from the following \href{https://www.dropbox.com/sh/1l9em6p0wyabwr3/AADhCpFvqjv6y7VbL5tgB3Iqa?dl=0}{link}.}

\subsection*{Methods}% Make the title.
\vspace{-4pt}
%\maketitle 
\subsubsection*{Paper selection}
\vspace{-10pt}
We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.  To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Table \ref{tabs1}) between 2004 and 2014. Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.  A separate set of 20 papers was also randomly selected, and was given to all observers to review independently in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.

%\centerline{
%%\hspace{1cm} 
%\parbox{11cm}{\noindent Table S1. The selected journals and their 2012 impact factors.}
%}
\vspace{-12pt}
%\begin{longtable}[h]
\begin{longtable}{lr}
%\centering
\caption{The selected journals and their 2012 impact factors.}\\
%\centering
%\begin{tabular}{lr}
  \hline
\textbf{Journal} & \textbf{Impact Factor} \\
\hline
\endhead 
  \hline
Ecology Letters & 17.95 \\ 
  Ecological Monographs & 8.09 \\ 
  Frontiers In Ecology And The Environment & 7.62 \\ 
  Global Ecology And Biogeography & 7.22 \\ 
  Global Change Biology & 6.91 \\ 
  Diversity And Distributions & 6.12 \\ 
  Methods In Ecology And Evolution & 5.92 \\ 
  Proceedings Of The Royal Society B-biological Sciences & 5.68 \\ 
  Journal Of Ecology & 5.43 \\ 
  Ecology & 5.17 \\ 
  Ecography & 5.12 \\ 
  Journal Of Biogeography & 4.86 \\ 
  Functional Ecology & 4.86 \\ 
  Journal Of Animal Ecology & 4.84 \\ 
  Journal Of Applied Ecology & 4.74 \\ 
  American Naturalist & 4.55 \\ 
  Conservation Biology & 4.36 \\ 
  Ecological Applications & 3.81 \\ 
  Biological Conservation & 3.79 \\ 
  Biogeosciences & 3.75 \\ 
  Bulletin Of The American Museum Of Natural History & 3.48 \\ 
  Biology Letters & 3.35 \\ 
  Oikos & 3.32 \\ 
  Behavioral Ecology & 3.22 \\ 
  Ecosystems & 3.17 \\ 
  Advances In Ecological Research & 3.08 \\ 
  Oecologia & 3.01 \\ 
  Landscape Ecology & 2.90 \\ 
  Agriculture Ecosystems \& Environment & 2.86 \\ 
  Ecological Economics & 2.85 \\ 
   \hline
%\end{tabular}
\label{tabs1}
\end{longtable}
% Place your abstract within the special {sciabstract} environment.

@@

We recorded six primary dimensions of ecological observations, three related to space and three related to time.  The space-related dimensions were resolution, extent, and actual extent. Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was defined as the summed area of all spatial replicates (i.e. N * resolution, where N is the number of spatial replicates, which we also recorded), or the area that ecologists observe in practice. In assessing spatial scales, our analysis only considered the Cartesian plane. We did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g. depth profiles in marine ecology). In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.

@@

A full definition of all dimensions and how they were recorded is contained within the answers to the list of questions below. This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, and adapted as necessary during the course of the study, in order to ensure methodological consistency (see next section).

@@

\subsubsection*{Scale FAQ}
\vspace{-10pt}
\noindent\underline{\textbf{General:}}
\def\Hitem{\item [Q\stepcounter{enumi}\arabic{enumi}.]} 
\def\Hsubitem{\item [Q\arabic{enumi}.\arabic{enumi}.]} 
\vspace{-0.5cm}
\begin{enumerate}
  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) book reviews; 3) model-only studies, particularly theoretical models that are not developed or tested against observations; 4) if they are experimental manipulations (but if a study has a mix of observational and experimental treatments, record the former but exclude the latter).

  \Hitem \emph{\textbf{What are the standard categories to be used for defining observational method?}} Define observational method according to the following categories: Remote sensing or other geographic data (e.g. non-remotely sensed GIS data), passive/automated data collection, field/direct observation, or paleo-reconstruction (tree rings, charcoal cores, etc).

@@

  \Hitem \emph{\textbf{What is sampling duration, and how do we record it?}} How long an individual observation of an individual point in space took to make. Sampling duration multiplied by the number of repeat observations is used to calculate \emph{actual duration} (see Q9). This value will often not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the sampling duration. For example, for a field based method with intensive plot methods, if not enough information is provided to estimate a plausible sampling duration, assign a token 1 day. For remote sensing observations you can assume one second (although the observations are effectively instantaneous).  
   \Hitem \emph{\textbf{What is duration, and how do we estimate it?}} The duration is the total period of time over which the phenomenon of interest was observed. More specifically, in the case of repeated observations, this is the total time that elapsed between the first and last observations at a given point in space (or of the same individual organism or community). For once-off (unreplicated) observations, this time is equivalent to the sampling duration. However, there may be cases where once-off observations have a longer duration than the sample duration. For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013}. The observations are therefore not strictly temporally replicated, but the authors control for year of collection in their subsequent analysis to avoid confounding effects.  In this case, we can consider the effective duration to be three years, as the temporal information is encoded in the analysis. 
  \Hitem \emph{\textbf{What is \emph{actual duration}, and how do we calculate it?}} The actual duration is the integral of sampling duration (see Q7), or the time spent making one observation of the phenomenon in question. To clarify, actual duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (duration), nor the integral of time spent in observing all spatial replicates (see Q10). 
  \Hitem \emph{\textbf{Should actual duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g. for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated in Q9, actual duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
  \Hitem \emph{\textbf{How do we record duration and actual duration when there are no repeat observations?}} In these cases, duration and actual duration are both equal to sampling duration. 
   \Hitem \emph{\textbf{How do I record interval in cases where the interval is inconsistent?  For example, in a study where observations were repeated in 1979, 1980, 1981, 1984, 2007, 2009?}} Find the time between each successive period, and then take the average of that (remember to convert to days!). If there are two or more sets of unevenly spaced days for each site/plot/measurement being taken in the study, then find the average interval for each, and average the averages.  
    \Hitem \emph{\textbf{How do you determine the interval for paleo-reconstructions?}} Use the minimum estimate for dating precision as the estimate of time between samples (e.g. 50 years in the study of European charcoal deposits \cite{molinari_exploring_2013}). 
    \Hitem \emph{\textbf{How do you determine the sample duration (our third time category) for paleo-reconstructions?}} Similar to the answer to the previous question, the sampling duration is also the same as the minimum estimate of dating precision. The logic behind this is that in such cases, where a sediment or tree core or similar measurement is being collected, this effectively represents a continuous ``observation'', and the value associated with the minimum (or other reported) interval is typically an average (or another summary statistic like the maximum) of the amount accumulated. 
    \Hitem \emph{\textbf{What about intervals and durations for instrument-collected, or automated sensor-collected, observations?}} These are often similar to the paleo-reconstruction case. Take the minimum temperature or daily rainfall recorded at a weather station, which require constant observation over 24 hours to report. In such cases, the interval and sampling duration are both 24 hours. On the other hand, automated logging systems often provide a series of high frequency observations that are collected instantaneously. In these cases the sampling duration should be a token one second (to keep consistent with remote sensing [Q7]), and the interval should be the period between successive instantaneous measurements. 
    \Hitem \emph{\textbf{How do you treat interval for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably interested in some season-dependent phenomenon (e.g. breeding behavior), the reported values should be pegged to the season, not averaged across the duration (the start and end dates of the study). So in this example it would be 10 days.

@@

@@

To ensure consistency within the main analysis, the lead author also reviewed each observer's results from their individual draw of papers and flagged values that appeared to deviate from the protocol for re-review by the observer. Revised values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.

@@

We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 378 observations over 1,000 iterations. For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting (from a uniform distribution) a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion. In certain cases, the perturbation resulted in values that were not physically possible (e.g. interval or actual duration longer than duration; actual extent larger than extent), in which case we adjusted the perturbed values for the dimension that should be the smallest to equal that of the one that should be largest (e.g. we set interval to equal the perturbed duration value). This perturbed set of observations provided a basis for estimating uncertainty within our extracted set of observational scales.

\subsubsection*{Scale metrics}
\vspace{-10pt}
In presenting results (Fig. 1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1,000 perturbed ensemble members in order to account for the large range in scale values. We calculated histograms for the four primary observational dimensions from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1,000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig. 1 main text).

To reveal the densities of observations within two dimensions (Fig. 2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations were concentrated. Bandwidths of varying resolutions were tested on kernel density estimates of sampling interval versus plot resolution to test how sensitive our results were to the bandwidth value (see Supplementary Results section).

@@

\noindent Where x is either extent or duration and y is its actual counterpart. We then evaluated how the magnitudes of difference varied with increasing values of extent/duration, using box plots to summarize decades within the same bins used to summarize the frequency distributions of the extent and duration of observations (Fig. 1B and D main text). Decades were calculated for each pair for all bootstrap replicates.

\subsubsection*{Trends in methods and scale}
To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period. To undertake the former assessment, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year. For the second analysis, we applied the same regression approach to the four primary dimensions (resolution, extent, interval, and duration), in order to assess whether there were any trends in observational scales.

@@

We assessed variability and consistency between observers along multiple dimensions. First, we assessed tghe degree to which observers agreed with respect to selecting or rejecting papers for review, using R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.

@@

Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table \ref{tabs2}). CV values were estimated from all six observers for resolution, actual extent, and interval, from five observers (all but Choi) for duration, and from three observers (Elsen, Estes, and Treuer) for extent. The smaller numbers for duration and extent reflect the fact that initial efforts were focused on estimating resolution, interval, and actual extent and actual duration. Fewer observers were available for subsequent efforts to assess duration and extent.

\subsubsection*{Between-observer coefficient of variation}
\vspace{-10pt}
We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table \ref{tabs2}) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.

\begin{table}[ht]
\captionsetup{width=1\linewidth}
\caption{The between-observer variability of estimates of the spatial and temporal scales of ecological observations reported within the set of 20 papers used for calibration. Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records.}
\vspace{-4 pt}
\begin{center}
\begin{tabular}{rR{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}}
\hline
& \multicolumn{3}{c}{\textbf{Spatial}} & \multicolumn{3}{c}{\textbf{Temporal}}\\\cline{2-7}
%Value & Resolution & Extent & \multirow{2}{*}{Actual} & Interval & Duration & Actual Duration \\
Value & Resolution & Extent & Actual Extent & Interval & Duration & Actual Duration \\
\hline
CV of overall mean & 50 & 50 & 78 & 49 & 36 & 109 \\
Mean of record-wise CV & 58 & 41 & 72 & 105 & 64 & 126 \\
\hline
\end{tabular}
\end{center}
\label{tabs2}
\end{table}%

@@

Fig. \ref{obshists} details the distributions of observations separated by observational method along the four primary dimensions. Fig. \ref{obskde} shows the densities of observations by observational method within the dimensional juxtapositions shown in Fig. 2 in the main text. Fig. \ref{obsdiff} shows the how the differences between actual extent and extent and actual duration and duration vary by observational type.

\subsubsection*{Trends in methods and scales}
\vspace{-10pt}
The results of the weighted linear regression applied to observational types by year (Fig. \ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (p$<$0.12 for regression of remote observations; p$<$0.18 for field observation regression).  Automated sensing methods showed no visible trend.

If this trend towards increasing use of remote sensing between 2004-2014 was real and not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average. We cannot attach much confidence to this projections given that it derived from a non-significant trend; however, some support for this projection lies within the regression analysis applied to extent values themselves, which showed an apparent increase in extent of 0.25 orders of magnitude per year (Fig. \ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This weak trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).  The other three dimensions did not show any clear year-on-year trends.
